---
title: "Session 1"
date: July 27, 2023
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Agenda

* Getting familiar with the RStudio IDE
* Using RStudio Projects
* Importing data from an Excel file
* Introducing `ggplot2` to create data visualizations

## Script

Here's the script from Session 1. You can download this and save it inside the R Project folder that you created for this workshop.

<!-- to add scripts -->

<!-- ```{r, include = FALSE, eval = FALSE} -->

<!-- # install and load these packages -->
<!-- devtools::install_github('yihui/xfun') -->

<!-- # embed_file() requires a few more packages -->
<!-- xfun::pkg_load2(c('base64enc', 'htmltools', 'mime')) -->
<!-- ``` -->

<!-- ```{r echo=FALSE} -->
<!-- # a embed single file -->
<!-- xfun::embed_file('session_01.R') -->
<!-- ``` -->



## Cheatsheets

* [RStudio IDE](https://rstudio.com/wp-content/uploads/2016/01/rstudio-IDE-cheatsheet.pdf)

* [ggplot2](https://rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf)


## Resources
A nice review of the RStudio IDE and R Projects:
https://www.youtube.com/watch?v=kfcX5DEMAp4

The introduction to data visualization we will cover today is based off of Chapter 3 (sections 3.1 - 3.3) in the R4DS textbook: https://r4ds.had.co.nz/data-visualisation.html

To learn about loess: https://en.wikipedia.org/wiki/Local_regression


## Homework Exercises

1. Create an excel file (in Microsoft Excel) with three columns of data. The first column should include both words and numbers and be 20 rows long. Read you file into R using the `read_xlsx()` function. Try three different ways to specify the sheet you want imported. Add missing values at random to the data, using *-99* to indicate missing values. Read in the data again, and make sure missing values are handled correctly.  

1. Assign the "correct" colors to the **colors** variable in the plot above. Take a look at this [link](https://www.statology.org/color-by-factor-ggplot2/) (Example 2) to figure out how to change the default color scheme.

    What did this example teach you about where the `aes()` command can be placed?

    Run `?scale_color_manual` in the R console (this will only work if the `ggplot2` package has been loaded). What happens? Find the paper talking about color vision deficiencies (and read it if you are interested).
    
    What does the `c()` function do?
    
1. Make all dots in the plot squares. Use the text book or the internet to figure out how it is done. What did you learn about shapes and colors?  

1. A colleague of yours, from one of the Scandinavian countries (you don't remember which, it could be any of them) has sent you a csv file. The data can be downloaded here. Read in the data using one of the functions from the `readr` package and make sure all columns are correctly read. You will have to open the data and figure out a way to identify where the data comes from (and use google to find the ICU locale language code). Then complete this code: 

```{r, eval=FALSE}
read_csv(file = ___,
         na = ___,
         locale = ___,
         col_types = cols(
           values = ___,
           date = ____
         ))
```


More information can be found in the R4DS book, chapters [11.3.4](https://r4ds.had.co.nz/data-import.html#readr-datetimes) and [11.4.2](https://r4ds.had.co.nz/data-import.html#problems). Make sure you include everything necessary to avoid all warning messages. 